---
title: "Variance and Std Deviation | Why divide by n-1?"
output:
 html_document:
    df_print: paged
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: haddock
    smart: false
editor_options: 
  chunk_output_type: inline
---


Hello World

Author: OA Lab, NWFSC
Title: JVariance and Standard Deviation
Date: 2021 (R document January 2021)


#1. Creating Petrol Dataset

```{r 1.0  Creating Petrol Dataset }

num.wk <- c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")

petrol.price <- c("48.50", "87.40", "19.98", 
               "59.74", "40.87", "105.51",
               "40.80", "23.10", "98.10",
               "60.54", "64.81", "48.01")

petrol.price <- as.numeric(petrol.price)

petrol.wk <- data.frame(num.wk, petrol.price)




mean.ptrl.wk <- mean(petrol.wk$petrol.price)
ptrl.wk.variance <- var(petrol.wk$petrol.price)
ptrl.wk.sd <- sd(petrol.wk$petrol.price)



```



```{r 1.1  Descriptive Statistics Petrol Dataset}

view(mean.ptrl.wk)
# 58.11333
view(ptrl.wk.variance)
# 748.0117
view(ptrl.wk.sd)
# 27.3498


```






```{r}

hist(petrol.wk$petrol.price) 

```



We bother with variance to describe the "spread of the data"
You can see the lowest observation to the highest observation
the average deviation from the mean (squared population mean)


Degrees of Freedom
how singular and unqiue each of our datapoints can be 
the number of independent pieces of information in the dataset

```{r}
dA1 <- c("1", "2", "3")

dA2 <- c("41", "59", "50")

dA2 <- as.numeric(dA2)

dA3 <- data.frame(dA1, dA2)
```

Effect Size 

Test Statistic: Observed Data - What We Expect if the Null if True
(over)
average variation

[Comparison of Distributions]: Users/katherinerovinski/GIT/2021.01.13_mixedeffectmodel/DistributionComparison.png

```{r}

```



average deviation  
























