---
title: "2021.01.27 Regression Learning"
output:
 html_document:
    df_print: paged
    code_folding: hide
    toc: TRUE
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: haddock
    smart: false
editor_options: 
  chunk_output_type: inline
---


Hello World

Author: OA Lab, NWFSC
Title: 2021.01.27 Regression Learning
Date: 2021 (R document January 2021)



```{r 0.0 Libraries , echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
#*********************************
##Libraries
#********************************* 
library(stringr)
library(tidyverse)
library(plyr)
library(nlme)
library(tidyr)
library(dbplyr)
library(dplyr)
library(purrr)
library(wql)
library(lubridate)
library(tibbletime)
library(arsenal)
library(compareDF)
#for graphing
library(ggplot2)
library(ggfortify)
library(stringr)
library(nlme)
library(RColorBrewer)
library(patchwork)
#statistical analysis
library(gdata)
library(rsq)
library(doBy)
library(lme4)
library(lmerTest)
#Rnotebooks 
library(gridExtra)
library(kableExtra)

```



#1. Creating Petrol Dataset

```{r 1.0  Creating Petrol Dataset }
# Bar Takings ... barTakings

takingsDate <- c("2021-06-03", 
                 "2021-06-10", 
                 "2021-06-17", 
                 "2021-06-24", 
                 "2021-07-01",
                 "2021-07-08", 
                 "2021-07-15", 
                 "2021-07-22", 
                 "2021-07-29")

takingsDollars <- c( 3213, 
                     2089, 
                     2253, 
                     1801, 
                     801, 
                     1943, 
                     1720,
                     1514, 
                     1017)

takingsTemp <- c(23, 
                 21,
                 25,
                 18,
                 13,
                 16,
                 13,
                 17,
                 12)

barTakings <- data.frame(takingsDate, takingsDollars, takingsTemp)


```




```{r}
# Simple Scatter Plot 

attach(barTakings)
plot(takingsDollars, takingsTemp, main="Scatterplot Example",
     xlab = "Takings Dollars", ylab = "Takings Temperature", pch=19)




```
Regression relationship is prehaps best defined as a line of best fit
Adding a fit line

```{r}

#Add fit lines
attach(barTakings)
plot(takingsDollars, takingsTemp, main="Scatterplot Example",
     xlab = "Takings Dollars", ylab = "Takings Temperature", pch=19)
abline(lm(takingsTemp~takingsDollars), col="red") # regression line y~x
lines(lowess(takingsTemp~takingsDollars), col="blue") # lowess fit line (x,y)


```


the r function abline can be used to add vertical, horizontal, or regression lines to a graph
simplified format of the abline() function
abline(a=NILL, b=NULL, v=NULL, ...)

a,b : single values specifying the intercept and the slope of the line
h : the y value(s) for the horizontal line(s)
v : the x value(s) for the vertical line(s)

LOWESS Locally Weighted Scatterplot Smoothing (LOESS- locally weighted smoothing)
popular tool used in regression analysis that creates a smooth line through a timeplot or scatter plot to help see relationships 



```{r}
#Add fit lines
attach(barTakings)
plot(takingsDollars, takingsTemp, main="Scatterplot Example",
     xlab = "Takings Dollars", ylab = "Takings Temperature", pch=19)
# abline(lm(takingsTemp~takingsDollars), col="red") # regression line y~x
# lines(lowess(takingsTemp~takingsDollars), col="blue") # lowess fit line (x,y)
takingsReg <- lm(takingsDollars ~ takingsTemp, data = barTakings)
coeff=coefficients(takingsReg)


eq = paste0("y =", round(coeff[2],1), "*x", round(coeff[1],1))
plot(takingsDollars, takingsTemp, main =eq)
abline(takingsReg, col="blue")


```

